---
title: Do email templates reduce my workload?
author: Luke McGuinness
date: '2021-07-11'
slug: working-with-emails
categories:
  - R
  - side-project
tags: 
excerpt: An attempted before-and-after analysis of email data in R.
math: no
meta: yes
---

```{r, include=FALSE}
knitr::opts_chunk$set(eval = FALSE, warning = F, message = F)
source(here::here("R","sidenote.R"))
```

As part of my maintenance of the `robvis` R package, I regularly receive emails from users of the Shiny app version of the tool.`r sidenote_link("https://mcguinlu.shinyapps.io/robvis/")`

One of the main issues I encounter when dealing with these emails is that the initial email does not usually contain all the information I need to help the user (e.g. what data file caused the issues), and so I end up replying requesting this basic information.

To try and head this issue off at the pass, I implemented a email template in March 2020, so that when users click the "Email me!" button on the Shiny app, an email is created and pre-populated with the following text:

``` {text}

This is a template to help me fix your issue as quickly as possible. Please make sure you:
- detail your issue, and where it occurred [e.g. data upload, generating graphs]
- attach the file you were trying to upload

Thanks,

Luke


```

I recently had the idea to explore whether this email template was actually helping in terms of my workload, by comparing the total number of emails in threads related to the `robvis` grouped by whether the thread occurred before or after the email template was implemented.`r sidenote("Commonly known as a \"before-and-after study\". See [here](https://www.bmj.com/content/349/bmj.g5074).")`

If the template does help, you would expect threads after March 2020 to contain a smaller number of emails, as it would remove the need for my initial email requesting the necessary data and the user's response to this request.

I recently have become quite good at labelling emails related to the different R packages I maintain and moving them to a dedicated Outlook subfolder. As such, I was able to follow a very helpful tutorial to export the `robvis` Outlook folder to a CSV file,`r sidenote_link("https://everydayanalytics.ca/2013/11/how-to-export-your-outlook-inbox-to-csv-for-data-analysis.html")` making sure to use the Outlook->Access->CSV route, so that the final dataset contained information on the date/time of each email. Once imported, I cleaned the data using the following chain:

```{r, eval = TRUE}

library(dplyr)

df <-
  read.csv("robvis_emails.csv",
           stringsAsFactors = F,
           header = F) %>%
  # Rename the v. unhelpful variable names and select the relevant ones
  rename(
    date = V15,
    text = V14,
    subject = V5,
    sender  = V6,
    subject_normalised = V19
  ) %>%
  select(date, sender, subject, subject_normalised, text) %>%
  # Remove the weird forwarding
  mutate(subject_normalised = stringr::str_replace(subject_normalised, "Antw: RE: ", "")) %>%
  # Convert dates and trim character variables
  mutate(date = as.Date(date, "%d/%m/%Y %H:%M:%S")) %>%
  mutate_if(is.character, stringr::str_trim) %>%
  # Group by sender/subject and get count and first date by group
  group_by(sender, subject_normalised) %>%
  summarise(Number = n(), date = min(date)) %>%
  ungroup()

```

One interesting thing I noticed is that, because of the way Outlook works, only the incoming emails (i.e. those sent to my email) in each thread are counted. As a result, several threads appear to only have a single email in them, but in reality, the users emailed me, I emailed them back and then never heard from them again. I _think_ this is because Outlook deals with incoming messages and sent messages differently, but I'm not 100% sure.

I then quickly plotted the data to have a look at the date range it covered.

`r tufte::margin_note("I used the <code>ggannotate</code> widget to help create the annotations for this plot.")`

``` {r, eval = TRUE, echo = FALSE}
library(ggplot2) 

ggplot(df) +
  aes(x = date) +
  aes(y = Number) +
  geom_point() +
  scale_y_continuous(breaks = seq(0, 10, by = 2)) +
  labs(title = "Number of incoming emails per thread") +
  geom_text(
    data = data.frame(
      x = as.Date("2021-01-01"),
      y = 9,
      label = "This long thread (N = 10) was with a colleague in\nBristol, and involved a discussion of a\ncustom plot type."
    ),
    mapping = aes(x = x,
                  y = y,
                  label = label),
    angle = 0L,
    lineheight = 1L,
    hjust = 0.5,
    vjust = 0.5,
    colour = "black",
    family = "sans",
    fontface = "plain",
    inherit.aes = FALSE,
    show.legend = FALSE
  ) +
  geom_curve(
    data = data.frame(
      x = as.Date("2020-10-03"),
      y = 8.9861196506449,
      xend = as.Date("2020-07-17"),
      yend = 9.85487558594912
    ),
    mapping = aes(
      x = x,
      y = y,
      xend = xend,
      yend = yend
    ),
    angle = 90L,
    colour = "black",
    curvature = -0.34,
    arrow = structure(list(
      angle = 30,
      length = structure(
        0.1,
        class = "unit",
        valid.unit = 2L,
        unit = "inches"
      ),
      ends = 2L,
      type = 2L
    ),
    class = "arrow"),
    inherit.aes = FALSE,
    show.legend = FALSE
  )


```

As it happens, I apparently only started adding emails to the `robvis` Outlook subfolder after July 2020. As the email template was implemented in March 2020, my plan to do a before-and-after comparison to see if it had any impact on the number of emails in each thread was scuppered. 

I could always go back through my emails from before July 2020 and categorise them appropriately, or alternatively, import my entire inbox and come up with a data cleaning plan that would extract the relevant emails. However, as this was intended as a quick exercise, I decided to park the analysis.

Still, it gave me an excuse to work with email data in R, so not a total loss!

