---
title: Formatting platform-agnostic file paths 
author: Luke McGuinness
date: '2021-02-02'
slug: pathformatr 
categories:
  - R
tags: []
math: no
meta: yes
draft: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
source(here::here("R","sidenote.R"))
```

## The problem 

I really love using the `here` package f

The advantages of this approach have been set out elsewhere,`r sidenote("See Jenny Bryan's [Ode to the here package](https://github.com/jennybc/here_here).")`. Under this setup, file paths are abstracted away from a seperator

```{r}
library(here)

here::here("data/2020/06/01/data.csv")

# should be

here::here("data","2020","06","01","data.csv")

```

When using this set-up you have two options. The first is to build the path up, element by element.

However, `r sidenote("For example, typing \"data/\" and pressing tab will bring up a drop-down list of all files within the data folder.")``

The second is to use RStudio's path auto-complete functionality, and then go back and 

However, this really doesn't play well with RStudio's auto-complete functionality, meaning that I either have to manually. 

## The solution

First issue was creating the functions to identify and then reformat the]

<!-- TODO Talk about how the package identifies and replaces -->

<!-- TODO Talk also about how this -->


### RELEASE THE REGEX!!

The hardest part of the project (but by far the most satisfying when it *finally* worked) was created the regular expression syntax to match and replace the slashes.

The worst by far was the backslash, where you need four backslashes to match a single literal one`r sidenote("See https://xkcd.com/1638/")`:

Apply the function to each match, to create the text to replace it with.


### Implementing the addin

Once you have your function (`format_path()`) ready to go, it's a case of creating a `inst/rstudio/addins.dcf` file in your package to "bind" it to an RStudio add-in.`r sidenote("See the fantastic [RStudio tutorial](https://rstudio.github.io/rstudioaddins/) on add-ins.")` I've reproduced the contents of the `pathformatr` `addins.dsf` file below: 


``` {}
Name: Split file path into quoted and comma-separated elements.
Description: Splits file paths in here() and file.path() calls into individual quoted and comma-separated segments, allowing users to autocomplete their paths using the RStudio file completion functionality and then quickly reformat to be serperator-agnostic.
Binding: format_path
Interactive: false
```


### Taking it a step further

Having implemented the functionality for a single highlighted file path, I wondered if I could use the same functions to reformat all `here()` or `file.path()` calls within the active document. 

An initial issue with `format_all_paths()` was that after running it, the cursor moves to the end of the active document. To combat this, the function grabs the line the cursor is on when the function is called and returns it to the start of this line once it is finished.

## Summary

Being completely honest, this saves me approximately 5 seconds each time I use it, so it's not going to revolutionise my workflow. However, it was an excuse to play around with the `rstudioapi` package.

However, hitting a keyboard short to implement an add-in that you wrote is a great feeling!