---
title: Blurbs and the Bechdel test
author: Luke McGuinness
date: '2021-04-12'
slug: do-blurbs-predict-whether-films-will-pass-the-bechdel-test
categories:
  - tidymodels
  - prediction
  - machine learning
  - R
  - side-project
tags: []
excerpt: Do promotional blurbs predict whether films will pass the Bechdel test?
math: no
meta: yes
draft: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, warning = FALSE)
source(here::here("R","sidenote.R"))
```

## Install relevant pacakges



## Load the data

For this analysis, I used a TidyTuesday dataset on the Bechdel test. `r sidenote_link("https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-03-09/readme.md")``

```{r}
library(tidyverse)
library(tidytext)

tuesdata <- tidytuesdayR::tt_load('2021-03-09')

movies <- tuesdata$movies %>%
  mutate(bechdel = binary) %>%
  select(title, plot, imdb_rating, bechdel) %>%
  filter(!is.na(plot) & !is.na(bechdel) & !is.na(imdb_rating)) %>%
  mutate(text = plot) %>%
  select(-plot)

# Number of words per movie blurb
movies %>%
  unnest_tokens(word, text) %>%
  count(title, name = "total_words") %>%
  ggplot(aes(total_words)) +
  geom_histogram(fill = "darkorange2", alpha = 0.8) +
  theme_minimal()


movies %>%
  ggplot(aes(y = imdb_rating, x = bechdel)) +
  geom_boxplot() +
  theme_minimal()
```

I also wanted to have a quick look at the balance between the two outcomes (Pass/Fail) to ensure the dataset was relatively well balanced.

```{r}
# Number of movies that pass/fail
movies %>%
  ggplot(aes(x = bechdel)) +
  geom_bar(stat = "count", fill = "darkorange2", alpha = 0.8) +
  theme_minimal()

prop.table(table(movies$bechdel))
```

It seems there are a few more movies that fail the test, but the split is not too far off 50/50.

### Build a model

First things first, split the data into training and testing dataset, using the default split of 3/4 training and 1/4 testing.

`r tufte::margin_note("I like using The Answer to the Ultimate Question of Life, the Universe, and Everything a a seed.")`

```{r}
library(tidymodels)

set.seed(42)
review_split <- initial_split(movies, strata = bechdel)
review_train <- training(review_split)
review_test <- testing(review_split)
```

Next, I made use of the amazing 

```{r}
library(textrecipes)
library(stopwords)

review_rec <- recipe(bechdel ~ text + imdb_rating, data = review_train) %>%
  step_tokenize(text) %>%
  step_stopwords(text) %>%
  step_tokenfilter(text, max_tokens = 500) %>%
  step_tfidf(text) %>%
  step_normalize(all_predictors())

review_prep <- prep(review_rec)

review_prep

```

###

```{r}
lasso_spec <- logistic_reg(penalty = tune(), mixture = 1) %>%
  set_engine("glmnet")

lasso_wf <- workflow() %>%
  add_recipe(review_rec) %>%
  add_model(lasso_spec)

lasso_wf
```


`r tufte::margin_note("You may need the development version of <code>recipes</code> for this to work, as per [this issue](https://github.com/tidymodels/tune/issues/192)")`

```{r, cache=TRUE}
lambda_grid <- grid_regular(penalty(), levels = 40)

set.seed(42)
review_folds <- bootstraps(review_train, strata = bechdel)
review_folds

doParallel::registerDoParallel()

set.seed(2020)
lasso_grid <- tune_grid(
  lasso_wf,
  resamples = review_folds,
  grid = lambda_grid,
  metrics = metric_set(roc_auc, ppv, npv)
)

lasso_grid %>%
  collect_metrics()
```



```{r}
best_auc <- lasso_grid %>%
  select_best("roc_auc")

best_auc

final_lasso <- finalize_workflow(lasso_wf, best_auc)

final_lasso

library(vip)

# final_lasso %>%
#   fit(review_train) %>%
#   pull_workflow_fit() %>%
#   vi(lambda = best_auc$penalty) %>%
#   group_by(Sign) %>%
#   arrange(desc(abs(Importance))) %>%
#   slice(1:20) %>%
#   ungroup() %>%
#   mutate(
#     Importance = abs(Importance),
#     Variable = str_remove(Variable, "tfidf_text_"),
#     Variable = fct_reorder(Variable, Importance)
#   ) %>%
#   ggplot(aes(x = Importance, y = Variable, fill = Sign)) +
#   geom_col(show.legend = FALSE) +
#   facet_wrap(~Sign, scales = "free_y") +
#   labs(y = NULL)


review_final <- last_fit(final_lasso, review_split)

review_final %>%
  collect_metrics()

review_final %>%
  collect_predictions() %>%
  conf_mat(bechdel, .pred_class)

```