---
title: Citation data in R
author: Luke McGuinness
date: '2021-02-16'
slug: citation-data-in-r
categories:
  - bibliographic-database
  - R
tags: []
excerpt: Put excerpt here
math: no
meta: yes
draft: true
---

```{r, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
source(here::here("R","sidenote.R"))
```

One of the packages I developed for creating publication quality risk-of-bias plots`r sidenote("[robvis](https://github.com/mcguinlu/robvis)")` has now been used in >100 systematic reviews. The functionality is available as a R`r sidenote("R")` package and as a Shiny web application.

In order to better understand the users of my software, I was keen to send out a short survey to those that had used the tool in a published paper.

## Extracting 

So how many papers have cited the `robvis` paper in Research Synthesis Methods:


```{r, class.output="fullwidth-output"}
library(rcrossref)

doi <- "10.1002/jrsm.1411"

citation_data <- cr_works(doi)

citation_data$data$is.referenced.by.count

```

However, the CrossRef API does not provide anyt 


